---
title: "COSHH Essentials: Qualitative Chemical Risk Assessment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{COSHH Essentials: Qualitative Chemical Risk Assessment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(expoquimR)
```

## Overview

COSHH Essentials (Control of Substances Hazardous to Health) is a qualitative
control-banding method developed by the UK Health and Safety Executive (HSE).
It is designed to help employers and occupational hygienists identify the
appropriate level of control for chemical substances without requiring
exposure measurements.

The method assigns substances to one of five **hazard groups** (A to E) based
on their hazard phrases (H or R), and combines this with the **quantity**
handled daily and the **volatility** (for liquids) or **dustiness** (for
solids) to produce a **risk level** (1 to 4) with associated recommended
control measures.

## Hazard groups

| Group | Hazard level | Typical H phrases |
|---|---|---|
| A | Low | H303, H315, H319, H336 |
| B | Moderate | H302, H312, H332, H371 |
| C | Significant | H301, H311, H314, H317, H331 |
| D | High | H300, H310, H330, H351, H360 |
| E | Very high (CMR) | H334, H340, H341, H350 |

Any substance with H or R phrases not listed in groups B–E is assigned to
group A by default (precautionary principle).

## Risk levels and control measures

| Risk level | Typical situation | Control approach |
|---|---|---|
| 1 | Low hazard, low quantity/volatility | General ventilation |
| 2 | Medium hazard or moderate quantity | Local exhaust ventilation |
| 3 | High hazard or large quantity | Containment or closed systems |
| 4 | CMR substances or very high hazard | Comply with CMR legislation; detailed assessment |

## Step-by-step example

### 1. Classify liquid volatility

For liquid substances, volatility is estimated from the boiling point and the
process temperature using the boundary lines defined in the COSHH Essentials
guide.

```{r volatility}
# Toluene: boiling point 111 °C, process temperature 20 °C
coshh_clasificar_volatilidad(t_ebullicion = 111, t_proceso = 20)
```

### 2. Assign the hazard group

```{r grade}
# Toluene: H315 (skin irritation), H336 (drowsiness/dizziness)
coshh_grado("H315, H336")
```

### 3. Obtain the risk level

```{r risk}
coshh_riesgo(grado = "A", cantidad = "Medium", volatilidad = "Medium")
```

### 4. Get the recommended control measures

```{r measures}
coshh_medidas(1)
```

### 5. Full assessment in one call

The high-level wrapper `coshh_evaluar()` chains all four steps:

```{r full}
coshh_evaluar(
  nombre       = "Toluene",
  frases       = "H315, H336",
  cantidad     = "Medium",
  es_liquido   = TRUE,
  t_ebullicion = 111,
  t_proceso    = 20
)
```

## Evaluating multiple substances

`coshh_evaluar()` is designed to work well with `lapply()` or `purrr::pmap()`
for batch evaluation:

```{r batch}
sustancias <- list(
  list(nombre = "Toluene",  frases = "H315, H336", cantidad = "Medium",
       es_liquido = TRUE,  t_ebullicion = 111, t_proceso = 20),
  list(nombre = "Xylene",   frases = "H315, H336", cantidad = "Large",
       es_liquido = TRUE,  t_ebullicion = 139, t_proceso = 20),
  list(nombre = "Cement",   frases = "H315",        cantidad = "Large",
       es_liquido = FALSE, pulverulencia = "High")
)

resultados <- do.call(rbind, lapply(sustancias, function(s) {
  do.call(coshh_evaluar, s)
}))

resultados
```

## From Excel (no coding required)

For users who prefer not to write R code, `expoquimR` includes a ready-to-fill
Excel template. Download it, fill in one row per substance, and pass the file
path to `coshh_desde_excel()`:

```{r excel, eval = FALSE}
# Access the built-in template
ruta <- system.file("plantillas", "plantilla_coshh.xlsx", package = "expoquimR")

# Or copy it to your working directory
file.copy(ruta, ".")

# Run the assessment
coshh_desde_excel("plantilla_coshh.xlsx")
```

## Language

All output labels are available in English (default) and Spanish:

```{r language}
expoquimr_lang("es")

coshh_evaluar(
  nombre       = "Tolueno",
  frases       = "H315, H336",
  cantidad     = "Mediana",
  es_liquido   = TRUE,
  t_ebullicion = 111,
  t_proceso    = 20
)

expoquimr_lang("en")  # reset to English
```

## Interactive application

For a point-and-click interface, launch the COSHH Shiny application:

```{r app, eval = FALSE}
run_coshh()
```

## References

- UK Health and Safety Executive (HSE). *COSHH Essentials: Easy steps to
  control chemicals.* Available at: https://www.hse.gov.uk/coshh/
- European Chemicals Agency (ECHA). *Understanding CLP — Hazard statements.*
  Available at: https://echa.europa.eu/
