---
title: "INRS Method: Qualitative Inhalation Risk Assessment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{INRS Method: Qualitative Inhalation Risk Assessment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(expoquimR)
```

## Overview

The INRS method is a qualitative control-banding approach for assessing the
risk of inhalation of chemical agents, developed by the French National
Research and Safety Institute (INRS). Unlike COSHH Essentials, which produces
a categorical risk level, the INRS method calculates a continuous **inhalation
risk score** as the product of five partial scores, each reflecting a different
dimension of the exposure situation.

The final score is compared against three risk bands to determine the
appropriate level of action.

## The five-factor cascade

| Factor | Source | Score range |
|---|---|---|
| Hazard potential | Hazard class (1–5) from H/R phrases, process or VLA | 1–10 000 |
| Quantity × Frequency | Exposure potential class (0–5) | 1–10 000 |
| Volatility / Dustiness | Physical state and conditions | 1, 10 or 100 |
| Process type | Degree of containment | 0.001–1 |
| Collective protection | Ventilation or enclosure | 0.001–10 |

The inhalation risk score is the product of all five:

```
Score = Hazard_potential × Volatility × Process × Protection × FC_VLA
```

## Risk bands

| Score | Risk band | Recommended action |
|---|---|---|
| ≤ 100 | Low | No modifications needed a priori |
| 101–1 000 | Moderate | Corrective measures or detailed assessment likely needed |
| > 1 000 | Very high | Immediate corrective action required |

## Step-by-step example

### 1. Hazard class

The hazard class (1–5) is determined from H phrases, R phrases, material and
process type, or VLA. The method explores from class 5 downwards and returns
the first match. If nothing matches classes 2–5, class 1 is assigned by
default.

```{r peligro}
# From H phrases
inrs_clase_peligro(frases_h = "H336")

# From VLA
inrs_clase_peligro(vla = 0.05)   # <= 0.1 mg/m³ → class 5
inrs_clase_peligro(vla = 50)     # > 10 mg/m³   → class 1 (default)
```

### 2. Quantity and frequency classes

```{r cantidad_frecuencia}
inrs_clase_cantidad(5, "l")          # 5 litres → class 2
inrs_clase_frecuencia(3, "horas")    # 3 hours/day → class 3
```

### 3. Exposure potential

```{r exposicion}
inrs_clase_exposicion_potencial("2", "3")   # → class 2
inrs_puntuacion_riesgo_potencial("2")        # → 10
```

### 4. Volatility (liquid)

Two approaches are available:

```{r volatilidad}
# From boiling point graph (official INRS Figure 2)
inrs_volatilidad_liquido_grafico(temperatura_uso = 20, punto_ebullicion = 111)

# From vapour pressure (Table 8 thresholds: 0.5 kPa / 25 kPa)
inrs_volatilidad_liquido_presion(presion_vapor = 5)
```

For solids, dustiness is classified by description:

```{r pulverulencia}
inrs_pulverulencia_solido("Polvo fino con poca dispersion visible")
```

### 5. Process and collective protection

```{r proceso_proteccion}
inrs_procedimiento("Abierto")
inrs_proteccion("Condiciones moderadas de dispersion")
```

### 6. VLA correction factor

```{r fc_vla}
inrs_fc_vla(vla = 50)    # VLA > 0.1 → FC = 1
inrs_fc_vla(vla = 0.05)  # VLA <= 0.1 → FC = 10
```

### 7. Final inhalation risk score

```{r riesgo}
inrs_riesgo_inhalacion(
  puntuacion_riesgo_potencial = 10,
  puntuacion_volatilidad      = 10,
  puntuacion_procedimiento    = 0.5,
  puntuacion_proteccion       = 0.7,
  fc_vla                      = 1
)
```

### 8. Risk characterisation

```{r caracterizacion}
inrs_caracterizacion(35)
inrs_caracterizacion(500)
inrs_caracterizacion(5000)
```

## Full assessment in one call

```{r full}
inrs_evaluar(
  nombre            = "Toluene",
  frases_h          = "H336",
  vla               = 50,
  cantidad_valor    = 5,
  cantidad_unidad   = "l",
  frecuencia_valor  = 3,
  frecuencia_unidad = "horas",
  tipo_sustancia    = "liquida",
  metodo_liquido    = "grafico",
  temperatura_uso   = 20,
  punto_ebullicion  = 111,
  procedimiento     = "Abierto",
  proteccion        = "Condiciones moderadas de dispersion"
)
```

## From Excel (no coding required)

```{r excel, eval = FALSE}
ruta <- system.file("plantillas", "plantilla_inrs.xlsx", package = "expoquimR")
inrs_desde_excel(ruta)
```

## Language

```{r language}
expoquimr_lang("es")
inrs_caracterizacion(35)
inrs_caracterizacion(5000)
expoquimr_lang("en")
```

## Interactive application

```{r app, eval = FALSE}
run_inrs()
```

The INRS Shiny application allows the user to evaluate up to 10 products in a
single session. Results are displayed in two panels: a summary table with the
key classes and the final risk score, and a collapsible detail panel showing
all intermediate scores for audit purposes.

## References

- INRS. *Evaluation du risque chimique par inhalation — Methode d'evaluation
  simplifiee du risque chimique.* ND 2233. Available at: https://www.inrs.fr/
