---
title: "UNE-EN 689: Quantitative Statistical Exposure Assessment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{UNE-EN 689: Quantitative Statistical Exposure Assessment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(expoquimR)
```

## Overview

UNE-EN 689 is a European standard that defines a two-stage statistical
procedure for comparing measured occupational exposure levels against
occupational exposure limits (OELs). It is designed to provide a statistically
rigorous and reproducible conformity decision based on a representative set of
exposure measurements collected over multiple working days.

The standard assumes that daily exposure (ED) values follow a **lognormal or
normal distribution** and uses a **one-sided upper tolerance limit**
(LSC\_{95,70}) to decide, with 95% confidence and 70% coverage, whether
exposure is below the OEL.

## The two-stage procedure

### Stage 1: Preliminary assessment (minimum 3 measurement days)

Each measurement day produces one or more samples. The daily exposure ED is
calculated as the time-weighted average over an 8-hour shift. The exposure
index IE = ED / OEL is then computed for each day.

Decision rules:

| Condition | Outcome |
|---|---|
| All IE < 0.1 | **CONFORMITY** — exposure is well below the OEL |
| Any IE > 1 | **NON-CONFORMITY** — at least one day exceeds the OEL |
| Any IE between 0.1 and 1 | **NO DECISION** — proceed to Stage 2 |

### Stage 2: Statistical assessment (minimum 6 measurement days total)

The full set of ED values (preliminary + additional days) is fitted to a
lognormal or normal distribution using the Shapiro-Wilk test. The one-sided
upper tolerance limit LSC\_{95,70} is then compared against the OEL via
the risk index UR. If UR ≥ UT (tabulated tolerance factor), conformity is
declared.

## Stage 1 example

```{r prelim}
# Five samples across three working days
datos <- data.frame(
  jornada       = c(1, 1, 2, 3, 3),
  concentracion = c(12, 8, 9, 5, 6),   # mg/m³
  tiempo        = c(4,  4, 8, 3, 5)    # hours
)

res_pre <- une689_evaluate_preliminary(datos, vla = 10)
res_pre$tabla_jornadas
res_pre$resultado
```

### Step-by-step

```{r prelim_steps}
# Daily exposure for day 1 (two samples)
une689_daily_exposure(concentracion = c(12, 8), tiempo = c(4, 4))

# Exposure index
une689_exposure_index(ed = 10, vla = 10)

# Classify conformity from a set of IE values
une689_classify_conformity(c(1.0, 0.9, 0.5625))
```

## Stage 2 example

When the preliminary assessment returns NO DECISION, additional measurement
days are collected and the full statistical assessment is performed on the
combined dataset.

```{r stat}
# Six ED values (3 preliminary + 3 additional)
eds <- c(10, 9, 5.625, 11, 8, 13)

res_est <- une689_evaluate_statistical(eds, vla = 10)

cat("Distribution:", res_est$tipo, "\n")
cat("MG =",  round(res_est$MG,  3), "\n")
cat("DSG =", round(res_est$DSG, 3), "\n")
cat("UT =",  res_est$ut, "\n")
cat("LSC(95,70) =", round(res_est$lsc, 3), "\n")
cat("UR =",  round(res_est$ur, 3), "\n")
cat("Result:", res_est$conformidad, "\n")
```

### Step-by-step

```{r stat_steps}
# Descriptive statistics
est <- une689_statistics(eds)
est$MG; est$DSG

# Normality and lognormality tests
test <- une689_normality_test(eds)
test$pval_lognormal

# Distribution type (lognormal has priority)
tipo <- une689_distribution_type(
  pval_normal    = test$pval_normal,
  pval_lognormal = test$pval_lognormal
)
tipo

# Tolerance factor UT for n = 6
une689_ut(6)

# LSC(95,70)
une689_lsc(tipo, ut = une689_ut(6), MG = est$MG, DSG = est$DSG)

# Risk index UR
une689_ur(tipo, vla = 10, MG = est$MG, DSG = est$DSG)

# Conformity decision
une689_statistical_conformity(ur = une689_ur(tipo, vla = 10,
                                               MG = est$MG, DSG = est$DSG),
                                ut = une689_ut(6))
```

## Periodic assessment (monitoring interval)

Once conformity has been established, the standard requires defining how
frequently measurements should be repeated. Two options are available:

```{r periodic}
# Option 1: MG or MA versus the OEL
une689_monitoring_interval_opt1(valor_referencia = res_est$MG, vla = 10)

# Option 2: LSC(95,70) versus the OEL
une689_monitoring_interval_opt2(lsc = res_est$lsc, vla = 10)
```

## Additive effects

When workers are simultaneously exposed to multiple agents affecting the same
target organ, the European standard requires that the **combined exposure
index** be evaluated:

```
IE_combined = IE_agent1 + IE_agent2 + ... + IE_agentN
```

Conformity requires IE_combined ≤ 1. For example, if toluene (IE = 0.20) and
xylene (IE = 0.30) both affect the central nervous system:

```{r additive}
ie_toluene <- 0.20
ie_xylene  <- 0.30
ie_combined <- ie_toluene + ie_xylene

cat("Combined IE:", ie_combined, "\n")
une689_classify_conformity(ie_combined)
```

Agents can appear in more than one additive group if they affect multiple
target organs. The `une689_from_excel()` function and the Shiny application
handle multiple independent additive groups automatically.

## From Excel (no coding required)

The UNE-EN 689 Excel template has three sheets:

- **Agents**: agent name and VLA, one row per agent.
- **Measurements**: one row per sample, with a `tipo` field (`pre` for
  preliminary days, `add` for additional days).
- **Additive_effects** (optional): groups of agents sharing a target organ.

```{r excel, eval = FALSE}
ruta <- system.file("plantillas", "plantilla_une689.xlsx", package = "expoquimR")
res <- une689_from_excel(ruta)

# Preliminary results and statistical assessment per agent
res$preliminar$Tolueno$resultado
res$preliminar$Tolueno$estadistica$conformidad

# Additive effects table
res$aditivo
```

## Language

```{r language}
expoquimr_lang("es")
une689_classify_conformity(c(0.02, 0.05))
une689_distribution_type(0.5, 0.5)
une689_monitoring_interval_opt1(0.5, vla = 10)
expoquimr_lang("en")
```

## Interactive application

```{r app, eval = FALSE}
run_une689()
```

The UNE-EN 689 Shiny application supports:

- Multiple chemical agents, each with its own tab and independent measurement
  days.
- Sequential workflow: preliminary assessment → automatic unlock of additional
  days on NO DECISION → statistical assessment → periodic assessment.
- Independent additive effect groups: users define which agents share a target
  organ, and agents can appear in more than one group.

## Methodological note

A known implementation error in some versions of this standard involves the
use of `all(IE < 0.1, na.rm = TRUE)` to test conformity. In R, `all()` on
an empty vector returns `TRUE`, which would incorrectly declare conformity
when no valid IE values are available. `expoquimR` corrects this by returning
`NA` from `une689_classify_conformity()` when the input contains no valid
IE values.

## References

- AENOR. *UNE-EN 689:2019. Workplace exposure — Measurement of exposure by
  inhalation to chemical agents — Strategy for testing compliance with
  occupational exposure limit values.* Madrid: AENOR.
- European Committee for Standardisation (CEN). *EN 689:2018.*
